---
title: 'Peer-graded Assignment: Prediction Assignment Writeup'
author: "Eric Fruelda"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prediction Assignment Writeup
This is a Prediction Assignment Writeup that was written as a requirement for Practical Machine Learning Course in Coursera.

## Getting the data
The data for this assignment came from this source: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har.

## Loading R Libraries that will be used
```{r}
library(caret)
```

## Loading and Cleaning the data
The data needs to be checked and explored for possible cleaning options.
```{r}
training <- read.csv("~/pml-training.csv", header=TRUE, na.strings=c("NA","#DIV/0!",""))
testing <- read.csv("~/pml-testing.csv", header=TRUE, na.strings=c("NA","#DIV/0!",""))
```

#### Checking the data
```{r eval=TRUE}
str(training)
head(training[,1:10])
tail(training[,1:10])
```

#### Removing NA columns
```{r}
training <- training[,colSums(is.na(training)) == 0]
testing <- testing[,colSums(is.na(testing)) == 0]
```

#### Removing X, ID, Timestamp and Window columns
```{r}
training <- training[c(8:60)]
testing <- testing[c(8:60)]
```

## Creating Partition for a Validation Dataset
The dataset is quite big so we can create partition for a validation dataset. The seed is set to 13563 for reproducibility.
```{r}
set.seed(13563)
subValid <- createDataPartition(y=training$classe, p=0.7, list=FALSE)
training <- training[subValid,]
validation <- training[-subValid,]
```

## Training the Data
#### Training and prediction using Random Forest Algorithm
```{r}
modFitRF <- train(classe ~., method = "rf", data = training)
predFitRF <- predict(modFitRF, validation)
```

#### Checking of Model on Validation Data using Confusion Matrix
```{r eval=TRUE}
confusionMatrix(predFitRF, as.factor(validation$classe))
```

## Prediction using Final Model on Testing Dataset
Because the model has resulted to 100% accuracy on the validation dataset, we can use the model as the final model to make predictions to our Testing data.

#### Plotting the decision tree
``` {r eval=TRUE}
plot(modFitRF$finalModel, main = "Random Forest Model")
```

#### Prediction on Testing Dataset
```{r eval=TRUE}
predFinal <- predict(modFitRF, testing)
predFinal
```